\PaperTitle

### 2020-05-06 paper/filter_bubbles.tex
nix conclusion

\xhdr{User Learning}
When a user consumes an item $n$ she learns the realized value for that item. In our model, we incorporate the idea that learning the value of item $n$ gives a user information about similar items. 
%This is drawn from recent empirical evidence in \cite{schulz2019structured} that consumers learn how to navigate large choice sets using similarity-based generalization. 
We assume that learning about the value of item $n$ reveals more about the value associated to items that are closer to it, which captures the idea that trying an item provides more information about similar items than about dissimilar ones.
\par
In order to have a well-defined notion of similarity we need to define a distance function between items, which we define as $d(n,m):=\min\{ \lvert m - n \rvert ,N - \lvert m - n \rvert \}$ where $m$ and $n$ are indices of items in $\mathcal{J}$. This distance function is not intended to model the intricacies of a realistic product space, but instead to provide a stylized product space to help us understand the effects of informational spillovers on user behavior. We consider that the entry of $n$-th row and the ($m$)-th column of $\Sigma_i$ is given by $\sigma_i^2 \rho^{d(n,m)}$, and that of $\Sigma$ is given by $\sigma^2 \rho^{d(n,m)}$. The scalar $\rho \in [0,1]$ therefore impacts the covariance structure: a higher $\rho$ implies that learning the utility of $n$ is more informative about products nearby. Informativeness, for any $\rho \in (0,1)$, is decreasing in distance. The particular distance function that we utilize leads to this covariance structure being simple, where the $(n,n+1)$-th entry in the covariance matrix is $\sigma^{2} \rho$, the $(n,n+2)$-th entry is $\sigma^{2} \rho^2$, etc. \footnote{This exponential decay correlation structure can be related to the tenet of case-based similarity of \cite{gilboa1995case} -- see \cite{billot2008axiomatization} for an axiomatization of exponential similarity.}
\par
The precise updating rule is as follows. Recall that at time $t$ the user's consumption history is given by $C_{i}^{t}$ and we denote the utility realizations of these items as $c_t$. We denote $\mu_t$ as the initial mean beliefs the user has over the items in $C_{i}^{t}$ and $\mu_{N-t}$ as the initial mean beliefs the user has over the remaining $N-t$ items, $\mathcal{J} \setminus C_{i}^{t}$. We partition the covariance matrix as follows:
\[ \Sigma =  \left( \begin{array}{cc}
\Sigma_{(N-t, N-t)} & \Sigma_{(N-t,t)} \\
\Sigma_{(t,N-t)} & \Sigma_{(t,t)}
\end{array} \right)
\]
Thus, after consuming the items in $C_{i}^{t}$ the resulting beliefs over the remaining items are given by $\mathcal{N}(\bar{\mu}, \bar{\Sigma})$ where $\bar{\mu}$ and $\bar{\Sigma}$ are as follows:
\begin{align*}
\bar{\mu} \mid c_t = \mu_{N-t} + \Sigma_{(N-t,t)} \Sigma_{(t,t)}^{-1}(c_t - \mu_t) \\
\bar{\Sigma} \mid c_t = \Sigma_{(N-t,N-t)} - \Sigma_{(N-t,t)} \Sigma_{(t,t)}^{-1} \Sigma_{(t,N-t)}
\end{align*}
\begin{figure}[t]
\includegraphics[width=.3\linewidth]{figures/Example-Bubbles.pdf}
\captionof{figure}{Illustrative Example}
\label{fig:illustrative_example}
\end{figure}

@@

Furthermore, our model shows the importance of understanding the underlying user preferences and the nature of the product space in the setting in which a recommender system is deployed since that will change the consumption patterns of users. For instance, understanding the degree of correlation of utilities and the level of risk-aversion become crucial in understanding how users will make consumption choices without recommendation and what kind of recommendation would be useful for users.
\par

\section{Conclusion}
We have studied a model of user decision-making in the context of markets where recommender systems are traditionally deployed in order to understand the consequences of recommender systems on consumption patterns of users. The key components of the model are that users make decisions under uncertainty and that consuming a good leads to informational spillovers about similar items. We have shown that this alone can generate consumption behavior that is consistent with filter bubble effects that have traditionally been associated with personalized recommender systems. This effect is weakened once we allow for recommendation, but not as much as if users have full information. We have further explored the relationship between consumption diversity and user welfare and found that they are negatively correlated, implying that striving for consumption diversity is not necessarily welfare-improving for users. Finally, we have shown that recommendation can lead to increases in user homogenization both relative to no recommendation and full information since it concentrates users' consumption in portions of the product space with the highest utility according to the limits of what the recommender can predict about the utility of the good.
\par 
We leave for future work exploring to what extent user behavior changes if we drop the myopia assumption, so that users themselves are engaging in explicit exploration. This likely interacts with recommendation in interesting ways as the recommendations may provide information that can complement or crowd-out user-led exploration. Even under the myopia assumption, understanding which recommendation system is optimal is an interesting and challenging problem. Further, our results point to an important consideration in the literature on incentivizing exploration which is that the exploration faced by the recommender itself likely interacts in interesting and important ways with the choice problem faced by the users.
\par 
Overall, understanding how users make choices in online platforms and how recommender systems impact these choices is important to understand. Understanding these issues allows us to better understand the broader social and economic impact of such systems as well as can provide a guide to improve the design of such systems.
% Bibliography
\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

>>>>

\xhdr{User Learning}
When a user consumes an item $n$ she learns the realized value for that item. In our model, we incorporate the idea that learning the value of item $n$ gives a user information about similar items. 
%This is drawn from recent empirical evidence in \cite{schulz2019structured} that consumers learn how to navigate large choice sets using similarity-based generalization. 
We assume that learning about the value of item $n$ reveals more about the value associated to items that are closer to it, which captures the idea that trying an item provides more information about similar items than about dissimilar ones.
\par
In order to have a well-defined notion of similarity we need to define a distance function between items, which we define as $d(n,m):=\min\{ \lvert m - n \rvert ,N - \lvert m - n \rvert \}$ where $m$ and $n$ are indices of items in $\mathcal{J}$. This distance function is not intended to model the intricacies of a realistic product space, but instead to provide a stylized product space to help us understand the effects of informational spillovers on user behavior. We consider that the entry of $n$-th row and the ($m$)-th column of $\Sigma_i$ is given by $\sigma_i^2 \rho^{d(n,m)}$, and that of $\Sigma$ is given by $\sigma^2 \rho^{d(n,m)}$. The scalar $\rho \in [0,1]$ therefore impacts the covariance structure: a higher $\rho$ implies that learning the utility of $n$ is more informative about products nearby. Informativeness, for any $\rho \in (0,1)$, is decreasing in distance. The particular distance function that we utilize leads to this covariance structure being simple, where the $(n,n+1)$-th entry in the covariance matrix is $\sigma^{2} \rho$, the $(n,n+2)$-th entry is $\sigma^{2} \rho^2$, etc. \footnote{This exponential decay correlation structure can be related to the tenet of case-based similarity of \cite{gilboa1995case} -- see \cite{billot2008axiomatization} for an axiomatization of exponential similarity.}
\par
The precise updating rule is as follows. Recall that at time $t$ the user's consumption history is given by $C_{i}^{t}$ and we denote the utility realizations of these items as $c_t$. We denote $\mu_t$ as the initial mean beliefs the user has over the items in $C_{i}^{t}$ and $\mu_{N-t}$ as the initial mean beliefs the user has over the remaining $N-t$ items, $\mathcal{J} \setminus C_{i}^{t}$. We partition the covariance matrix as follows:
\[ \Sigma =  \left( \begin{array}{cc}
\Sigma_{(N-t, N-t)} & \Sigma_{(N-t,t)} \\
\Sigma_{(t,N-t)} & \Sigma_{(t,t)}
\end{array} \right)
\]
Thus, after consuming the items in $C_{i}^{t}$ the resulting beliefs over the remaining items are given by $\mathcal{N}(\bar{\mu}, \bar{\Sigma})$ where $\bar{\mu}$ and $\bar{\Sigma}$ are as follows:
\begin{align*}
\bar{\mu} \mid c_t = \mu_{N-t} + \Sigma_{(N-t,t)} \Sigma_{(t,t)}^{-1}(c_t - \mu_t) \\
\bar{\Sigma} \mid c_t = \Sigma_{(N-t,N-t)} - \Sigma_{(N-t,t)} \Sigma_{(t,t)}^{-1} \Sigma_{(t,N-t)}
\end{align*}

@@

Furthermore, our model shows the importance of understanding the underlying user preferences and the nature of the product space in the setting in which a recommender system is deployed since that will change the consumption patterns of users. For instance, understanding the degree of correlation of utilities and the level of risk-aversion become crucial in understanding how users will make consumption choices without recommendation and what kind of recommendation would be useful for users.

% Bibliography
\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}
